---
title: "Estaimting Mean from Quantiles"
author: "Jonathon Hirschi"
date: "June 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(pacman)) install.packages("pacman")

pacman::p_load(tidyverse, pander, wesanderson)

```

## Intro

Accuracy of different methods for estimating the sample mean using the quantiles are explored.

[Bland method (2014)](https://pdfs.semanticscholar.org/59dd/526a4335850fcb364c92bb6f4eb879fb6e59.pdf) uses sample size, 3 quartiles, min, and max.

Using the minimum and maximum seems dubious in practice, and ill-defined in theory.

## Data Simulation

Data is simulated from the normal distribution, gamma (shape = 1, scale = 2), and gamma (shape = 2, scale = 2). The distributions are plotted below (means displayed are population level).

100 replications of 1,000 observations. Data is organized in a `tibble`, utilizing the `purrr::map` function.

```{r sim}

# Parameters
N = 500
nreps = 100

simDF <- tibble(
  distr = factor(c('normal', 'gamma12', 'gamma22')),
  FUN = list(normal = function(x) rnorm(x), gamma12 = function(x) rgamma(x, shape = 1, scale = 2), gamma22 = function(x) rgamma(x, shape = 2, scale = 2))
)

sim_data <- function(FUN, N, nreps) return(matrix(FUN(N*nreps), nrow = N, ncol = nreps))

set.seed(555)
simDF <- simDF %>% 
  mutate(data = map(FUN, function(f) sim_data(f, N, nreps)))

simDF

```

```{r viz, echo = F}
pal <- wes_palette('Darjeeling1', nlevels(simDF$distr), 'discrete')
a <- 0.001
p1 <- ggplot(data = data.frame(x = qnorm(c(a, 1-a))), aes(x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 1), size = 1,
                color = pal[simDF$distr == 'normal']) + 
  geom_vline(xintercept = 0, col = pal[simDF$distr == 'normal'], linetype = 'dashed', size = 1) +
  annotate('text', x = 0, y = .18, label = 'Mean = 0', 
           col = pal[simDF$distr == 'normal'], angle = -90, vjust = -.7) +
  labs(y = 'Density', x = 'X', title = 'Normal Distribution')

p1 + theme_bw()


p2 <- ggplot(data = data.frame(x = qgamma(c(a, 1-a), shape = 2, scale = 2)), aes(x)) +
  stat_function(fun = dgamma, n = 101, args = list(shape = 2, scale = 2), size = 1, col = pal[simDF$distr == 'gamma22'])+
  geom_vline(xintercept = 4, col = pal[simDF$distr == 'gamma22'], linetype = 'dashed', size = 1) +
  annotate('text', x = 4, y = .08, label = 'Mean = 4', 
           col = pal[simDF$distr == 'gamma22'], angle = -90, vjust = -.7) +
  labs(y = 'Density', x = 'X', title = 'Gamma Distribution (shape = 2, scale = 2)')

p2 + theme_bw()

p3 <- ggplot(data = data.frame(x = qgamma(c(a, 1-a), shape = 1, scale = 2)), aes(x)) +
  stat_function(fun = dgamma, n = 101, args = list(shape = 1, scale = 2), size = 1, color = pal[simDF$distr == 'gamma12']) +
  geom_vline(xintercept = 2, col = pal[simDF$distr == 'gamma12'], linetype = 'dashed', size = 1) +
  annotate('text', x = 2, y = .3, label = 'Mean = 2', 
           col = pal[simDF$distr == 'gamma12'], angle = -90, vjust = -.7) +
  labs(y = 'Density', x = 'X', title = 'Gamma Distribution (shape = 1, scale = 2)')

p3 + theme_bw()
```

## Estimator

For sample mean $\bar{y}$, estimators are:

$$\hat{\bar{y}} = mean(p10 + Median + p90)$$
$$\hat{\bar{y}} = mean(p1 + Median + p99)$$

```{r validate}
estimators <- tibble(
  name = c('quant', 'decile', 'p10p90'),
  vals = list('quant' = c(0, .25, .5, .75, 1), 'decile' = seq(0, 1, by = .1),
              'p10p90' = c(.1, .5, .9))
)

str(estimators)

est_wrap <- function(vec, est) {
  mean(quantile(vec, est))
}

col_est <- function(data, est) {
  apply(data, 2, function(cl) est_wrap(cl, est))
} 




simDF <- simDF %>% 
  mutate(samp_mean = map(data, colMeans),
         quant_est = map(data, function(d) col_est(d, estimators$vals$quant)),
         dec_est = map(data, function(d) col_est(d, estimators$vals$decile)),
         p10p90_est = map(data, function(d) col_est(d, estimators$vals$p10p90)))

rmse <- function(x, y) {
  sqrt(mean((x - y)^2))
}

simDF <- simDF %>% 
  mutate(
    quant_est_acc = map2_dbl(quant_est, samp_mean, rmse),
    dec_est_acc = map2_dbl(dec_est, samp_mean, rmse),
    p10p90_est_acc = map2_dbl(p10p90_est, samp_mean, rmse)
  )
simDF
```



